Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-11-15 response_to_reviewers.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Major comments from Reviewer 1

> Overall, I thought the authors were successful in introducing the big ideas in sports
analytics at a general and non-technical level, and many interesting aspects of the literature
were discussed. However, I think the second part of the title ”Statistical Tools for Their
Investigation” was not properly explored. Several packages are mentioned throughout, and
a brief sketch of a possible analysis workflow is given in Section 5, but I think it lacks sufficient
detail. Beyond that, I think there are some extraneous figures and possible improvements
with the overall flow.

Our revision includes a significant bolstering of Sections 4, 5, and 6. In particular, we have added a subsection on Elo ratings, added a subsection on baseball tools, expanded and linked the discussion of graphical tools, replaced some weak figures with more relevant substitutes, and added supporting references.

@@

> Pages 3,4, 5: The use of historical tables and data is powerful.  However, I worry about the quality of the images particularly Figure 2.

We share this concern. We are also uncertain about any licensing restrictions. We have left the images as is for now, and will wait direction from Wiley.

> Page 6, line 8:  "the effect may be overstated" is somewhat ambiguous.  Is it the

We have clarified our language on this subject. The reason for the comment was related to the increased resolution of player tracking data. We have made this reasoning explicit and added citations to support it.

### 2022-11-15 sports_analytics.Rmd
close #34
close #42
references for sports betting; #29

@lopez2020bigger presents an introduction to NFL tracking data, and examines 4th down behavior as an example of the type of problem that can be more thoroughly studied with the increase in granularity of the tracking data over traditional NFL data.  
In the past, when looking at down and distance data to study whether NFL coaches are making good decisions about whether to "got for it" or punt on 4th down, the distance data is only a rounded approximation of the true distance "to go" (i.e. 1 yard, 1 foot, and 1 inch will all be recorded as 4th and 1. 
In fact, anything up to 2 yard will recorded as 4th and 1 @lopezDM). However, a coach on the field during a game will be able to clearly see the difference between 1 inch and 1 yard, and this information will factor into their decision making. 
With tracking data, the "to go" distance can be much more accurately assessed and therefore evaluation of 4th down coaching decisions can now account for this "extra" information that is available to a coach on the field of play, but not recorded in traditional NFL data. 
Many past analyses of the decision to go for it or not on 4th down conclude that coached in the NFL are too conservative in their decisions making. 
@lopez2020bigger also concludes that coaches are too conservative on 4th down decision making, but the effect may be overstated due to the way in which to go yardage has been recorded in the past.

@@

# Sports Betting
@LopezBaumerMatthews2018 uses betting market data to [describe paper]

@Boulier2017

@@

@@

Analytical work in sports requires facility with an ever-changing set of computational tools for working with data. 
Sources of authoritative data about sports are myriad, and are too numerous to list here. 
Software tools for sports analytics are similarly numerous. 
For R, we maintain a [CRAN Task View for Sports Analytics](https://CRAN.R-project.org/view=SportsAnalytics) that catalogs R packages published on the Comprehensive R Archive Network (CRAN) and organizes them by sport [@baumer2022ctv].
Table \ref{tab:ctv} provides an overview of the currently available sport-specific CRAN packages.
Further, a more general collection of software tools is being curated by the SportsDataverse initiative [@gilani2022sportsdataverse].

>>>>

@lopez2020bigger presents an introduction to NFL tracking data, and examines 4th down behavior as an example of the type of problem that can be more thoroughly studied with the increase in granularity of the tracking data over traditional NFL data.
In the past, when looking at down and distance data to study whether NFL coaches are making good decisions about whether to "got for it" or punt on 4th down, the distance data is only a rounded approximation of the true distance "to go" (i.e. 1 yard, 1 foot, and 1 inch will all be recorded as 4th and 1. 
In fact, anything up to 2 yard will recorded as 4th and 1 [@lopezDM]. However, a coach on the field during a game will be able to clearly see the difference between 1 inch and 1 yard, and this information will factor into their decision making. 
With tracking data, the "to go" distance can be much more accurately assessed and therefore evaluation of 4th down coaching decisions can now account for this "extra" information that is available to a coach on the field of play, but not recorded in traditional NFL data. 
Many past analyses of the decision to go for it or not on 4th down conclude that coached in the NFL are too conservative in their decisions making. 
@lopez2020bigger also concludes that coaches are too conservative on 4th down decision making, but the effect may be overstated due to the way in which to go yardage has been recorded in the past.

@@

# Sports betting

While @Lopez2018 uses betting market data to model team strengths, it does not directly address strategies for betting or inefficiencies in betting markets. 
Work such as @lacey1990estimation, @Boulier2006,

@Boulier2006 that there is insufficient evidence to conclude that betting markets were inefficient from 1994--2000.
@boulier2003predicting test several predictors for NFL games and find that the betting market is the best predictor.
@Buttrey2016 demonstrate a method that produces a positive ROI.
@gandar1988testing conclude that stats test are too weak to reject the hypothesis of rationality.  But they do reject rationalist with ``economic tests". 
@lacey1990estimation : Finds the market to be efficient, though they did find some profitable betting rules over the period of time between 1984 and 1986

## Betting pools

@@

@breiter1997play

@clair2007optimal

@kaplan2001march

@metrick1996march

@niemi2008contrarian

@sauer1998economics

@imbrogno2022computing
Description: The average number of winning survivor pool entries is computed for the 2019 National Football League (NFL) season. The probability of having to share the winning pot, given realistic expectations on game outcomes, is also shown.

@nichols2014impact conclude that the impact of travel is not completely incorporated into the betting markets, but that any effect is too small to find any profitable advantage.

@PaulWeinbach2014 Market efficiency and behavioral biases in the wnba betting market

@soebbing2013gamblers

@spann2009sports NBA bettors believe that tanking is occurring.

@@

Analytical work in sports requires facility with an ever-changing set of computational tools for working with data. 
Sources of authoritative data about sports are myriad, and are too numerous to list here. 
Software tools for sports analytics are similarly numerous. 
For R, we maintain a [CRAN Task View for Sports Analytics](https://CRAN.R-project.org/view=SportsAnalytics) that catalogs R packages published on the Comprehensive R Archive Network (CRAN) and organizes them by sport [@baumer2022ctv].
Table \ref{tab:ctv} provides an overview of the currently available sport-specific CRAN packages.
@casals2022systematic offers a systematic review of sport-related packages on CRAN.
Further, a more general collection of software tools is being curated by the SportsDataverse initiative [@gilani2022sportsdataverse].
